---
title: "Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(RCurl)
library(lubridate)
library(dplyr)
library(knitr)
library(stargazer)
library(pastecs)
library(zoo)
library(forecast)
library(xts)


```



```{r read_data}

data11 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2011%20MVT%20Location%20Q.csv")
data.2011 <- read.csv(text=data11)

data12 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2012%20MVT%20Location%20Q.csv")
data.2012 <- read.csv(text=data12)

data13 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2013%20MVT%20Location%20Q.csv")
data.2013 <- read.csv(text=data13)

data14 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2014%20MVT%20Location%20Q.csv")
data.2014 <- read.csv(text=data14)

data15 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2015%20MVT%20Location%20Q.csv")
data.2015 <- read.csv(text=data15)

FastMover <- read.csv(text=getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2015%20Fast%20Mover%20Review.csv"))
# FastMover2016 <- read.csv("2016 Fast Mover Review.csv", stringsAsFactors = FALSE)

```


```{r cleaning_ data}

data.2011.clean <- subset(data.2011,select = c(Material,Fast,Item,Posting,MvT,Quantity,uom,Local.Value,Group.Value))
data.2012.clean <- subset(data.2012,select = c(Material,Fast,Item,Posting,MvT,Quantity,uom,Local.Value,Group.Value))
data.2013.clean <- subset(data.2013,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))
data.2014.clean <- subset(data.2014,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))
data.2015.clean <- subset(data.2015,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))

colnames(data.2011.clean) <- colnames(data.2014.clean)
colnames(data.2012.clean) <- colnames(data.2014.clean)
colnames(data.2013.clean) <- colnames(data.2014.clean)
colnames(data.2015.clean) <- colnames(data.2014.clean)


```


```{r transforming.data}

data.all <- rbind(data.2011.clean,data.2012.clean)
data.all <- rbind(data.all,data.2013.clean)
data.all <- rbind(data.all,data.2014.clean)

#Subset data.all to obtain list of all fast items
FastData <- data.all[data.all$Fast=="Yes",]
FastData2015 <- data.2015.clean[data.2015.clean$Fast=="Yes",]

#remove the Fast Data column as it is no longer needed
FastData$Fast <- NULL
FastData2015$Fast <- NULL

#converting the Posting date to a standard date notation
# and setting quantity, Local Value and Group value to numeric data type

FastData$Posting <- gsub("\\.","-",FastData$Posting)
FastData$Posting <- dmy(FastData$Posting)

FastData$Quantity <- as.numeric(gsub(",","",FastData$Quantity))
FastData$Local.Value <- as.numeric(gsub(",","",FastData$Local.Value))
FastData$Group.Value <- as.numeric(gsub(",","",FastData$Group.Value))

FastData$Issue.Receipt <- ifelse(FastData$Quantity>=0,1,0)

# Repeated for the test data set - FastData2015

FastData2015$Posting <- gsub("\\.","-",FastData2015$Posting)
FastData2015$Posting <- dmy(FastData2015$Posting)

FastData2015$Quantity <- as.numeric(gsub("\\,","",FastData2015$Quantity))
FastData2015$Local.Value <- as.numeric(gsub("\\,","",FastData2015$Local.Value))
FastData2015$Group.Value <- as.numeric(gsub("\\,","",FastData2015$Group.Value))

FastData2015$Issue.Receipt <- ifelse(FastData2015$Quantity>=0,1,0)

```

#column list
```{r }
str(FastData)
str(FastMover)


```


```{r  missing.values}
# Check number of missing values in FastData
paste0("Missing Values by Variables :")
sapply(FastData, function(x) {sum(is.na(x))})

# Check number of missing values in FastData2015
paste0("Missing Values by Variables :")
sapply(FastData2015, function(x) {sum(is.na(x))})

```

#Comparing Mvt Material variable (2011-2014) to Fast Mover (2015) Material variable
```{r  }
paste0("Number of Transactional Materials :" , length(unique(FastData$Material)))
paste0("Number of Fast Mover Materials :" , length(unique(FastMover$Material)))
MvtMaterials <- as.data.frame(sort(unique(FastData$Material)))
FastMoverMaterials <- as.data.frame(sort(FastMover$Material))
colnames(MvtMaterials) <- "Materials"
colnames(FastMoverMaterials) <- "Materials"

# Fast Mover review Materials that did not match with materials in the transactional report
anti_join(FastMoverMaterials,MvtMaterials)

# Materials in the transactional report that did not match with Fast Movers
anti_join(MvtMaterials,FastMoverMaterials)

```
#Comparing Mvt Material variable (2015) to Fast Mover (2015) Material variable
```{r  }
paste0("Number of Transactional Materials :" , length(unique(FastData2015$Material)))
paste0("Number of Fast Mover Materials :" , length(unique(FastMover$Material)))
MvtMaterials2015 <- as.data.frame(sort(unique(FastData2015$Material)))
FastMoverMaterials <- as.data.frame(sort(FastMover$Material))
colnames(MvtMaterials2015) <- "Materials"
colnames(FastMoverMaterials) <- "Materials"

# Fast Mover review Materials that did not match with materials in the transactional report
anti_join(FastMoverMaterials,MvtMaterials2015)

# Materials in the transactional report that did not match with Fast Movers
anti_join(MvtMaterials2015,FastMoverMaterials)

```





#Analysing 2011-2014 Mvt variable; #Subset FastData for all issued inventory
```{r }
table(FastData$MvT)
FastIssue <- filter(FastData, MvT %in% c("201", "221", "261", "291"))

# Convert all issued quantities to postive numbers for ease of analysis
FastIssue$Quantity <- FastIssue$Quantity * -1

head(FastIssue)

```

#Analysing 2015 Mvt variable; #Subset FastData2015 for all issued inventory
```{r }
table(FastData2015$MvT)
FastIssue2015 <- filter(FastData2015, MvT %in% c("201", "221", "261", "291"))

# Convert all issued quantities to postive numbers for ease of analysis
FastIssue2015$Quantity <- FastIssue2015$Quantity * -1

head(FastIssue2015)

```


#Calculating transactions per month for 2011 - 2014  
```{r transactions.per.month}

#Create Month and Year variables in FastIssue data
FastIssue$Month <- month(FastIssue$Posting)
FastIssue$Year  <- year(FastIssue$Posting)
head(FastIssue)

#### Data Summaries
#Unique nummber of materials in FastIssue
length(unique(FastIssue$Material))

#Number of times a material was issued by year and month
month.summ <- 
        FastIssue %>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n())


#Total number of months a material was issued (in descending order)
item.summ <- 
        month.summ %>%
        group_by(Material) %>%
        summarise(month.count = n()) %>%
        arrange(desc(month.count))
        
# Total number of times material appeared in the data set (in descending order)
data.item <-
        FastIssue %>%
        group_by(Material) %>%
        summarise(count = n()) %>%
        arrange(desc(count))

# A data set consisting of a join of all monthly summaries
count.summ <- month.summ %>%
                  left_join(item.summ, by="Material") %>%
                  left_join(data.item, by ="Material") %>%
                  arrange(desc(month.count),Material,Year,Month)

# Summary of material quantities issued by month and year
quantity.summ <-
        FastIssue %>%
        group_by(Material,Year,Month) %>%
        summarise(Quantities = sum(Quantity)) %>%
        arrange(desc(Quantities))


# Full data set containing summary of quantities issued by Year and month
SumQuantity <- left_join(count.summ,quantity.summ,by=c("Material","Year","Month"))

# Dropping materials not stored in the Warehouse

SumQuantity2 <- filter(SumQuantity,!(Material %in% c("495848","495885","495893","578916",
                                                     "578926","578931","593649","599965",
                                                     "700148","720330", "724656", "724659")))
# Write item.sum and count.summ 

write.csv(SumQuantity2, "All Count.csv")

```

#Calculating transactions per month for 2015. Prepare test data set
```{r transactions.per.month}

#Create Month and Year variables in FastIssue data
FastIssue2015$Month <- month(FastIssue2015$Posting)
FastIssue2015$Year  <- year(FastIssue2015$Posting)
head(FastIssue2015)

#### Preparing test data set by aggregating issuing transactions for the year
#Unique nummber of materials in FastIssue2015
length(unique(FastIssue2015$Material))

#Number of times a material was issued by year
Year.summ2015 <- 
        FastIssue2015 %>%  
        group_by(Material,Year) %>%
        summarise(Quantities = sum(Quantity))

# Dropping materials not stored in the Warehouse

test2015 <- filter(Year.summ2015,!(Material %in% c("495848","495885","495893","578916",
                                                     "578926","578931","593649","599965",
                                                     "700148","720330", "724656", "724659")))
# Write item.sum and count.summ 

write.csv(test2015, "test2015.csv")

```

#Prepare training data set
```{r}
                      ############################
                      ###   Training Data      ###
                      ############################

#### Create sample time-Series data sets. One for a material transacted over 48 months######
#### and another for a material transacted over less than 48 months######

# Subset data for Material 487999 which has transactions covering the complete 48 months
# This approach is to be used for all materials that have 48 months of transactions
SumQuantity3 <- SumQuantity2[,-c(5,6)]
MatNo487999 <- filter(SumQuantity2[,-c(4:6)],(Material=="487999"))
# # Combine separate Year and Month columns into single Date Column
# MatNo487999$Date <- as.yearmon(paste(MatNo487999$Year, MatNo500076$Month), "%Y %m")

#convert MatNo487999 data frame into time-series data
Mat487999.ts <- ts(MatNo487999[,c(4)], start=c(2011, 1), end=c(2014, 12), frequency=12)
Mat487999.ts
plot(Mat487999.ts)


# Monthly summary of 2015 test set
FastIssue2015.month.summ <- 
        FastIssue2015 %>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n(),Quantities = sum(Quantity))

FastIssue2015.month.summ$Date <- as.yearmon(paste(FastIssue2015.month.summ$Year,
                                                  FastIssue2015.month.summ$Month), "%Y %m")


#Subset data for Material 500076 which has transactions covering 42 months
# This approach is to be used for all materials that have less than 48 months of transactions
MatNo500076 <- filter(SumQuantity2[,-c(4:6)],(Material=="500076"))

# Create a new date column by combining the Year and Month columns into single Date Column
MatNo500076$Date <- as.yearmon(paste(MatNo500076$Year, MatNo500076$Month), "%Y %m")

#convert MatNo500076 into a regular time-series data using the zoo package
# Create zoo object
zMatNo500076 <- zoo(MatNo500076$Quantities, as.yearmon(MatNo500076$Date))
# create a zero width series consisting of 48 months
gMatNo500076 <- as.yearmon(seq(as.numeric(as.yearmon("2011-01")),
                as.numeric(as.yearmon("2014-12")), 1/12))

# merge zMatNo500076 with the zero width series to fill in the missing months with 0s
zzMatNo500076 <- merge(zMatNo500076, zoo(x=NULL,gMatNo500076), fill = 0)
# convert to time series
Mat500076.ts <- as.ts(zzMatNo500076)
Mat500076.ts
plot(Mat500076.ts)

#### Reconfiguring the training data set for batch forecasting

batch <- SumQuantity2[,-c(4:6)]
batch$Date <- as.yearmon(paste(batch$Year, batch$Month), "%Y %m")
batch.1 <- batch[, c("Material","Date","Quantities","Year","Month")]
batch.1 <- batch.1[,1:3]

# # batch <- FastIssue[,-c(2,4,6:11)]
colnames(batch.1) <- c("MaterialID","Date","Quantity")

# split data into groups by material id
dataByMaterialId <- split(batch.1, batch.1$MaterialID)

# create an xts object for each id
# xts_list <- lapply(dataByMaterialId, function(id) {
#   names <- list(NULL, paste0("Material.", id$MaterialID[1]))
#   xts(id$Quantity, as.Date(id$Date, "%Y-%d-%m"), dimnames = names)
# })

xts_list <- lapply(dataByMaterialId, function(id) {
  names <- list(NULL, paste0("Material.", id$MaterialID[1]))
  xts(id$Quantity, as.yearmon(id$Date), dimnames = names)
})

# use do.call + merge to combine all your xts objects into one object
# This code creates a data set with Month and Year in the first column and
# the Materials summarized by monthly quantity in the other columns
train <- do.call(merge, c(xts_list, fill = 0))

write.csv(train, "train.csv")

```

```{r}
generate.forecast <- function(material,train.start,test.start,test.end)
{
  
  data.train.material <- batch[batch$Material== material,]
  data.test.material  <- FastIssue2015.month.summ[FastIssue2015.month.summ$Material==material,]
  
  train.ts <- ts(data.train.material$Quantities, start=train.start, frequency = 12)
  test.ts <- ts(data.test.material$Quantities, start=test.start, frequency = 12)
  
  
  #model 1 :  Just use overall average
  
  all.month.mean <- mean(train.ts)
  pred.ts <- rep(all.month.mean,12)
  fcst.all.avg.rmse <- sqrt(mean(test.ts - pred.ts)^2)
  fcst.avg <- ts(pred.ts,frequency = 12,start = test.start)
  
  
  
  #model 2 : Use monthly average
  
  month.mean.df <- 
    data.train.material %>%
    group_by(Month) %>%
    summarise(month.mean = mean(issue.qty))
  
  
  pred.ts <- ts(month.mean.df$month.mean,frequency = 12,start = test.start)
  fcst.month.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  fcst.mon.avg <- pred.ts
  
  
  #Model 3 ( Auto.Arima)
  
  fit3 <- auto.arima(train.ts, stepwise=FALSE, approximation=FALSE)
  fcst.arima.1 <- forecast(fit3,h=length(test.ts))
  
  pred.ts <-  fcst.arima.1$mean
  arima.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  
  
  #Model 4 stl
  
  fit4 <- stl(train.ts,s.window = 21,t.window = 21)
  fcst.stl <- forecast(fit4, method="naive",h=length(test.ts))
  
  
  pred.ts <-  fcst.stl$mean
  stl.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  
  
  #model 5 nnet
  fit4 <- nnetar(train.ts)
  fcst4 <- forecast(fit4,h=length(test.ts))
  
  pred.ts <- fcst4$mean
  nnet.rmse <- sqrt(mean(test.ts - pred.ts)^2)
  
  
  #model 5.1
  fit5.1 <- hw(train.ts,damped=TRUE)
  fcst5.1 <- forecast(fit5.1,h=length(test.ts))
  sum(fcst5.1$mean)
  
  fit5.2 <- hw(train.ts)
  fcst5.2 <- forecast(fit5.2,h=length(test.ts))
  sum(fcst5.2$mean)
  
  
  
  fit5.3 <- ets(train.ts)
  fcst5.3 <- forecast(fit5.3,h=length(test.ts))
  sum(fcst5.3$mean)
  
  
  fit5.4 <- ets(train.ts,damped=TRUE)
  fcst5.4 <- forecast(fit5.4,h=length(test.ts))
  sum(fcst5.4$mean)
  
  
  
  ############################################
  
  test.length <- length(test.ts)
  rolling.yearly.sum <- rollapply(train.ts,12,sum,align="right")
  
  # model 6 forecast monthly data rolled into yearly and forecasted -- Arima
  fit6 <- auto.arima(rolling.yearly.sum)
  fcst6 <- forecast(fit6,h=12)
  
  # model 7 forecast monthly data rolled into yearly and forecasted -- STL
  fit7 <- stl(rolling.yearly.sum,s.window = 12,t.window=12)
  fcst7 <- forecast(fit7,h=12)
  
  # model 8 forecast monthly data rolled into yearly and forecasted -- HW
  fit8 <- hw(rolling.yearly.sum,seasonal="additive")
  fcst8 <- forecast(fit8,h=12)
  fcst8$mean
  
  
  # model 9  forecast monthly data rolled into yearly and forecasted -- ETS
  fit9 <- ets(rolling.yearly.sum,damped = TRUE)
  fcst9 <- forecast(fit9,h=12)
  fcst9$mean
  
  
  # model 10  forecast monthly data rolled into yearly and forecasted -- HW
  fit10 <- hw(rolling.yearly.sum,damped = TRUE)
  fcst10 <- forecast(fit10,h=12)
  fcst10$mean
  
  
  ###printing all forecast
  
  plot(test.ts)
  lines(test.ts,col=2)
  lines(window(fcst.avg,start =test.start, end = test.end),col=3)
  lines(window(fcst.mon.avg,start =test.start, end = test.end),col=4)
  lines(window(fcst.arima.1$mean,start =test.start, end = test.end),col=5)
  lines(window(fcst.stl$mean,start =test.start, end = test.end),col=6)
  lines(window(fcst4$mean,start =test.start, end = test.end),col=7)
  lines(window(fcst5.3$mean,start =test.start, end = test.end),col=8)
  legend("topright",lty=1, pch=2, col=2:8, cex = 0.5,
         c("Test Fcst","Average" ,"Monthly Average","Arima","STL","NNetar","ETS"))
  
  
  
  
  results.df <- data.frame( "Material" = c(material),
                            "Avg.fcst.rmse" =c(fcst.all.avg.rmse),
                            "Avg.monthly.fcst.rmse" = c(fcst.month.rmse),
                            "Arima.rmse" = c(arima.rmse),
                            "STL.rmse" = c(stl.rmse),
                            "Nnet.rmse" = c(nnet.rmse),
                            "Total.test.fcst" = c(sum(test.ts)),
                            "Total.avg.fcst" = c(sum(fcst.avg)),
                            "Total.month.avg.fcst" = c(sum(fcst.mon.avg)),
                            "Total.Arima.fcst" = c(sum(fcst.arima.1$mean)),
                            "Total.STL.fcst" = c(sum(fcst.stl$mean)),
                            "Total.Nnet.fcst" = c(sum(fcst4$mean)),
                            "Total.HW.Damped.Fcst" = c(sum(fcst5.1$mean)),
                            "Total.HW.Fcst" = c(sum(fcst5.2$mean)),
                            "Total.ets.Fcst" = c(sum(fcst5.3$mean)),
                            "Total.ets.damped.Fcst" = c(sum(fcst5.4$mean)),
                            "Total.Batch.Arima.Fcst" =  c(fcst6$mean[test.length]),
                            "Total.Batch.STL.Fcst"  = c(fcst7$mean[test.length]),
                            "Total.Batch.HW.Fcst" = c(fcst8$mean[test.length]),
                            "Total.Batch.ETS.Fcst" = c(fcst9$mean[test.length]),
                            "Total.Batch.HW.Damped.Fcst" = c(fcst10$mean[test.length]))
  
  
  
  results.df
  
}
```

```{r calling.generate.forecast}

res <- generate.forecast(487999,c(2011,1),c(2015,1),c(2015,12))
melt(res)

```