---
title: "Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(RCurl)
library(lubridate)
library(dplyr)
library(knitr)
library(stargazer)
library(pastecs)
library(zoo)
library(forecast)
library(xts)
library(reshape2)


```



```{r read_data}

data11 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2011%20MVT%20Location%20Q.csv")
data.2011 <- read.csv(text=data11)

data12 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2012%20MVT%20Location%20Q.csv")
data.2012 <- read.csv(text=data12)

data13 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2013%20MVT%20Location%20Q.csv")
data.2013 <- read.csv(text=data13)

data14 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2014%20MVT%20Location%20Q.csv")
data.2014 <- read.csv(text=data14)

data15 <- getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2015%20MVT%20Location%20Q.csv")
data.2015 <- read.csv(text=data15)

FastMover <- read.csv(text=getURL("https://raw.githubusercontent.com/Udiee/XPN-Inventory-Analytics/master/Data/2015%20Fast%20Mover%20Review.csv"))
# FastMover2016 <- read.csv("2016 Fast Mover Review.csv", stringsAsFactors = FALSE)

```


```{r cleaning_ data}

data.2011.clean <- subset(data.2011,select = c(Material,Fast,Item,Posting,MvT,Quantity,uom,Local.Value,Group.Value))
data.2012.clean <- subset(data.2012,select = c(Material,Fast,Item,Posting,MvT,Quantity,uom,Local.Value,Group.Value))
data.2013.clean <- subset(data.2013,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))
data.2014.clean <- subset(data.2014,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))
data.2015.clean <- subset(data.2015,select = c(Material,Fast,Item,Posting,MvT,Quantity,UOM,Local.Value,Group.Value))

colnames(data.2011.clean) <- colnames(data.2014.clean)
colnames(data.2012.clean) <- colnames(data.2014.clean)
colnames(data.2013.clean) <- colnames(data.2014.clean)
colnames(data.2015.clean) <- colnames(data.2014.clean)


```


```{r transforming.data}

data.all <- rbind(data.2011.clean,data.2012.clean)
data.all <- rbind(data.all,data.2013.clean)
data.all <- rbind(data.all,data.2014.clean)

#Subset data.all to obtain list of all fast items
FastData <- data.all[data.all$Fast=="Yes",]
FastData2015 <- data.2015.clean[data.2015.clean$Fast=="Yes",]

#remove the Fast Data column as it is no longer needed
FastData$Fast <- NULL
FastData2015$Fast <- NULL

#converting the Posting date to a standard date notation
# and setting quantity, Local Value and Group value to numeric data type

FastData$Posting <- gsub("\\.","-",FastData$Posting)
FastData$Posting <- dmy(FastData$Posting)

FastData$Quantity <- as.numeric(gsub(",","",FastData$Quantity))
FastData$Local.Value <- as.numeric(gsub(",","",FastData$Local.Value))
FastData$Group.Value <- as.numeric(gsub(",","",FastData$Group.Value))

FastData$Issue.Receipt <- ifelse(FastData$Quantity>=0,1,0)

# Repeated for the test data set - FastData2015

FastData2015$Posting <- gsub("\\.","-",FastData2015$Posting)
FastData2015$Posting <- dmy(FastData2015$Posting)

FastData2015$Quantity <- as.numeric(gsub("\\,","",FastData2015$Quantity))
FastData2015$Local.Value <- as.numeric(gsub("\\,","",FastData2015$Local.Value))
FastData2015$Group.Value <- as.numeric(gsub("\\,","",FastData2015$Group.Value))

FastData2015$Issue.Receipt <- ifelse(FastData2015$Quantity>=0,1,0)

```

#column list
```{r }
str(FastData)
str(FastMover)


```


```{r  missing.values}
# Check number of missing values in FastData
paste0("Missing Values by Variables :")
sapply(FastData, function(x) {sum(is.na(x))})

# Check number of missing values in FastData2015
paste0("Missing Values by Variables :")
sapply(FastData2015, function(x) {sum(is.na(x))})

```

#Comparing Mvt Material variable (2011-2014) to Fast Mover (2015) Material variable
```{r  }
paste0("Number of Transactional Materials :" , length(unique(FastData$Material)))
paste0("Number of Fast Mover Materials :" , length(unique(FastMover$Material)))
MvtMaterials <- as.data.frame(sort(unique(FastData$Material)))
FastMoverMaterials <- as.data.frame(sort(FastMover$Material))
colnames(MvtMaterials) <- "Materials"
colnames(FastMoverMaterials) <- "Materials"

# Fast Mover review Materials that did not match with materials in the transactional report
anti_join(FastMoverMaterials,MvtMaterials)

# Materials in the transactional report that did not match with Fast Movers
anti_join(MvtMaterials,FastMoverMaterials)

```
#Comparing Mvt Material variable (2015) to Fast Mover (2015) Material variable
```{r  }
paste0("Number of Transactional Materials :" , length(unique(FastData2015$Material)))
paste0("Number of Fast Mover Materials :" , length(unique(FastMover$Material)))
MvtMaterials2015 <- as.data.frame(sort(unique(FastData2015$Material)))
FastMoverMaterials <- as.data.frame(sort(unique(FastMover$Material)))
colnames(MvtMaterials2015) <- "Materials"
colnames(FastMoverMaterials) <- "Materials"

# Fast Mover review Materials that did not match with materials in the transactional report
anti_join(FastMoverMaterials,MvtMaterials2015)

# Materials in the transactional report that did not match with Fast Movers
anti_join(MvtMaterials2015,FastMoverMaterials)

```


#Analysing 2011-2014 Mvt variable; #Subset FastData for all issued inventory
```{r }
table(FastData$MvT)
FastIssue <- filter(FastData, MvT %in% c("201", "221", "261", "291"))

# Convert all issued quantities to postive numbers for ease of analysis
FastIssue$Quantity <- FastIssue$Quantity * -1

head(FastIssue)


```

#Analysing 2015 Mvt variable; Subset FastData2015 for all issued inventory
```{r }
table(FastData2015$MvT)
FastIssue2015 <- filter(FastData2015, MvT %in% c("201", "221", "261", "291"))

# Convert all issued quantities to postive numbers for ease of analysis
FastIssue2015$Quantity <- FastIssue2015$Quantity * -1

head(FastIssue2015)

```


#Calculating transactions per month for 2011 - 2014  
```{r transactions.per.month}

#Prepare FastIssue train for Modeling
FastIssue2 <- FastIssue[,-c(2,4,6:11)]
FastMonthlyQuant <- expand.grid(unique(FastIssue2$Material),
                                as.Date(min(FastIssue2$Posting):max(FastIssue2$Posting), 
                                        origin = '1970-01-01'))
names(FastMonthlyQuant) <- c('Material', 'Posting')

FastMonthlyQuant <- merge(FastMonthlyQuant, FastIssue2, by = c('Material', 'Posting'), all.x = T)
FastMonthlyQuant$Quantity[is.na(FastMonthlyQuant$Quantity)] <- 0

#Create Month and Year variables in FastIssue data
FastMonthlyQuant$Month <- month(FastMonthlyQuant$Posting)
FastMonthlyQuant$Year  <- year(FastMonthlyQuant$Posting)
head(FastMonthlyQuant)

#### Data Summaries
#Unique nummber of materials in FastIssue
length(unique(FastMonthlyQuant$Material))

#Number of times a material was issued by year and month
month.summ <- 
        FastMonthlyQuant %>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n())


#Total number of months a material was issued (in descending order)
item.summ <- 
        month.summ %>%
        group_by(Material) %>%
        summarise(month.count = n()) %>%
        arrange(desc(month.count))
        
# Total number of times material appeared in the data set (in descending order)
data.item <-
        FastMonthlyQuant %>%
        group_by(Material) %>%
        summarise(count = n()) %>%
        arrange(desc(count))

# A data set consisting of a join of all monthly summaries
count.summ <- month.summ %>%
                  left_join(item.summ, by="Material") %>%
                  left_join(data.item, by ="Material") %>%
                  arrange(desc(month.count),Material,Year,Month)

# Summary of material quantities issued by month and year
quantity.summ <-
        FastMonthlyQuant %>%
        group_by(Material,Year,Month) %>%
        summarise(Quantities = sum(Quantity)) %>%
        arrange(desc(Quantities))


# Full data set containing summary of quantities issued by Year and month
SumQuantity <- left_join(count.summ,quantity.summ,by=c("Material","Year","Month"))

# Dropping materials not stored in the Warehouse

train <- filter(SumQuantity,!(Material %in% c("495848","495885","495893","578916",
                                                     "578926","578931","593649","599965",
                                                     "700148","720330", "724656", "724659")))
# Write item.sum and count.summ 

write.csv(train, "train.csv")

```

#Calculating transactions per month for 2015. Prepare test data set
```{r transactions.per.month}

#Preapare FastIssue2015 for Model testing
FastIssue2015.2 <- FastIssue2015[,-c(2,4,6:9)]
FastMonthlyQuant2015 <- expand.grid(unique(FastIssue2015.2$Material),
                                as.Date(min(FastIssue2015.2$Posting):max(FastIssue2015.2$Posting), 
                                        origin = '1970-01-01'))
names(FastMonthlyQuant2015) <- c('Material', 'Posting')

FastMonthlyQuant2015 <- merge(FastMonthlyQuant2015, FastIssue2015.2, 
                              by = c('Material', 'Posting'), all.x = T)
FastMonthlyQuant2015$Quantity[is.na(FastMonthlyQuant2015$Quantity)] <- 0


#Create Month and Year variables in FastMonthlyQuant2015 data
FastMonthlyQuant2015$Month <- month(FastMonthlyQuant2015$Posting)
FastMonthlyQuant2015$Year  <- year(FastMonthlyQuant2015$Posting)
head(FastMonthlyQuant2015)

#### Preparing test data set by aggregating issuing transactions for the year
#Unique nummber of materials in FastIssue2015
length(unique(FastMonthlyQuant2015$Material))

#Number of times a material was issued by year and month
month.summ2015 <- 
        FastMonthlyQuant2015 %>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n())


#Total number of months a material was issued (in descending order)
item.summ2015 <- 
        month.summ2015 %>%
        group_by(Material) %>%
        summarise(month.count = n()) %>%
        arrange(desc(month.count))
        
# Total number of times material appeared in the data set (in descending order)
data.item2015 <-
        FastMonthlyQuant2015 %>%
        group_by(Material) %>%
        summarise(count = n()) %>%
        arrange(desc(count))

# A data set consisting of a join of all monthly summaries
count.summ2015 <- month.summ2015 %>%
                  left_join(item.summ2015, by="Material") %>%
                  left_join(data.item2015, by ="Material") %>%
                  arrange(desc(month.count),Material,Year,Month)

# Summary of material quantities issued by month and year
quantity.summ2015 <-
        FastMonthlyQuant2015 %>%
        group_by(Material,Year,Month) %>%
        summarise(Quantities = sum(Quantity)) %>%
        arrange(desc(Quantities))


# Full data set containing summary of quantities issued by Year and month
SumQuantity2015 <- left_join(count.summ2015,quantity.summ2015,by=c("Material","Year","Month"))

# Dropping materials not stored in the Warehouse

test2015 <- filter(SumQuantity2015,!(Material %in% c("495848","495885","495893","578916",
                                                     "578926","578931","593649","599965",
                                                     "700148","720330", "724656", "724659")))
# Write item.sum and count.summ 

write.csv(test2015, "test2015.csv")

```

#Matching up the training set and test set
```{r}
### Check to ensure materials in training set match with materials in the test set ####
paste0("Number of Train Materials :" , length(unique(train$Material)))
paste0("Number of Test Materials :" , length(unique(test2015$Material)))
train.materials <- as.data.frame(sort(unique(train$Material)))
test.materials <- as.data.frame(sort(unique(test2015$Material)))
colnames(train.materials) <- "Material"
colnames(test.materials) <- "Material"

# Test Materials that are not in the training set
TestNotTrain <- anti_join(test.materials,train.materials)

# Materials in the training set that are not in the test set
TrainNotTest <- anti_join(train.materials,test.materials)

#Subsetting training and test set so that all the materials in the training set are the same 
# materials that are in the test set

a <- subset(train, !(train$Material %in% TrainNotTest$Material))
b <- subset(test2015, !(test2015$Material %in% TestNotTrain$Material))

# Checking that the materials are now the same
paste0("Number of Train Materials :" , length(unique(a$Material)))
paste0("Number of Test Materials :" , length(unique(b$Material)))

# Creating the final traininng and test set. Removing unwanted variables
trainFinal <- a[,-c(4:6)]
test2015Final <- b[,-c(4:6)]

# Data sets for data visualization
# Compute correct issue counts per month
FastIssue.summ <- FastIssue %>%
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n())

FastIssue.summ <- filter(FastIssue.summ,!(Material %in% c("495848","495885","495893","578916",
                                                          "578926","578931","593649","599965",
                                                          "700148","720330", "724656", "724659")))

# Merge the correct issue count with the training data
trainFinalMonthly <- trainFinal %>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n(),Quantity = sum(Quantities))

trainFinalMonthly$Date <- as.yearmon(paste(trainFinalMonthly$Year,
                                                  trainFinalMonthly$Month), "%Y%m")

trainFinalIssue <- left_join(trainFinalMonthly,FastIssue.summ,by=c("Material","Year","Month"))
trainFinalIssue$issue.count.y[is.na(trainFinalIssue$issue.count.y)] <- 0
trainFinalIssue$issue.count.x <-  NULL
colnames(trainFinalIssue)[6] <- "Issue.count"

write.csv(trainFinalIssue, "trainFinalIssue.csv")

# Test data for Visualization
test2015FinalMonthly <- test2015Final%>%  
        group_by(Material,Year,Month) %>%
        summarise(issue.count = n(),Quantity = sum(Quantities))

test2015FinalMonthly$Date <- as.yearmon(paste(test2015FinalMonthly$Year,
                                                  test2015FinalMonthly$Month), "%Y%m")
test2015FinalMonthly$issue.count <- NULL

write.csv(test2015FinalMonthly, "test2015FinalMonthly.csv")

```

#Building the Forecast Models
``` {r}
###### Generating forecast models ##############
#generate.forecast <- function(material,train.start,test.start,test.end)
generate.forecast <- function(material)  
{
  
  data.train.material <- trainFinal[trainFinal$Material== material,]
  data.test.material  <- test2015Final[test2015Final$Material==material,]
  
  ###added by sj --start
  
  start.year.train <- min(data.train.material$Year)
  start.month.train <- min(data.train.material[data.train.material$Year == start.year.train,]$Month)

  start.year.test <- min(data.test.material$Year)
  start.month.test <- min(data.test.material[data.test.material$Year == start.year.test,]$Month)


  end.year.test <- max(data.test.material$Year)
  end.month.test <- max(data.test.material[data.test.material$Year == end.year.test,]$Month)

  train.start <- c(start.year.train,start.month.train)
  test.start <- c(start.year.test,start.month.test)
  test.end <- c(end.year.test,end.month.test)

  #### end added by sj
  
  
  
  train.ts <- ts(data.train.material$Quantities, start=train.start, frequency = 12)
  test.ts <- ts(data.test.material$Quantities, start=test.start, frequency = 12)
  
  
  #model 1 :  Just use overall average
  
  all.month.mean <- mean(train.ts)
  pred.ts <- rep(all.month.mean,12)
  fcst.all.avg.rmse <- sqrt(mean(test.ts - pred.ts)^2)
  fcst.avg <- ts(pred.ts,frequency = 12,start = test.start)
  
  
  #model 2 : Use monthly average
  
  month.mean.df <- 
    data.train.material %>%
    group_by(Month) %>%
    summarise(month.mean = mean(Quantities))
  
  
  pred.ts <- ts(month.mean.df$month.mean,frequency = 12,start = test.start)
  fcst.month.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  fcst.mon.avg <- pred.ts
  
  
  #Model 3 ( Auto.Arima)
  
  fit3 <- auto.arima(train.ts, stepwise=FALSE, approximation=FALSE)
  fcst.arima.1 <- forecast(fit3,h=length(test.ts))
  
  pred.ts <-  fcst.arima.1$mean
  arima.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  
  
  #Model 4 stl
  
  fit4 <- stl(train.ts,s.window = 21,t.window = 21)
  fcst.stl <- forecast(fit4, method="naive",h=length(test.ts))
  
  
  pred.ts <-  fcst.stl$mean
  stl.rmse <- sqrt(mean(test.ts - pred.ts)^2) 
  
  
  #model 5 nnet
  fit4 <- nnetar(train.ts)
  fcst4 <- forecast(fit4,h=length(test.ts))
  
  pred.ts <- fcst4$mean
  nnet.rmse <- sqrt(mean(test.ts - pred.ts)^2)
  
  
  #model 5.1
  fit5.1 <- hw(train.ts,damped=TRUE)
  fcst5.1 <- forecast(fit5.1,h=length(test.ts))
  sum(fcst5.1$mean)
  
  fit5.2 <- hw(train.ts)
  fcst5.2 <- forecast(fit5.2,h=length(test.ts))
  sum(fcst5.2$mean)
  
  
  
  fit5.3 <- ets(train.ts)
  fcst5.3 <- forecast(fit5.3,h=length(test.ts))
  sum(fcst5.3$mean)
  
  
  fit5.4 <- ets(train.ts,damped=TRUE)
  fcst5.4 <- forecast(fit5.4,h=length(test.ts))
  sum(fcst5.4$mean)
  
  
  
  ############################################
  
  test.length <- length(test.ts)
  rolling.yearly.sum <- rollapply(train.ts,12,sum,align="right")
  
  # model 6 forecast monthly data rolled into yearly and forecasted -- Arima
  fit6 <- auto.arima(rolling.yearly.sum)
  fcst6 <- forecast(fit6,h=12)
  
  # model 7 forecast monthly data rolled into yearly and forecasted -- STL
  fit7 <- stl(rolling.yearly.sum,s.window = 12,t.window=12)
  fcst7 <- forecast(fit7,h=12)
  
  # model 8 forecast monthly data rolled into yearly and forecasted -- HW
  fit8 <- hw(rolling.yearly.sum,seasonal="additive")
  fcst8 <- forecast(fit8,h=12)
  fcst8$mean
  
  
  # model 9  forecast monthly data rolled into yearly and forecasted -- ETS
  fit9 <- ets(rolling.yearly.sum,damped = TRUE)
  fcst9 <- forecast(fit9,h=12)
  fcst9$mean
  
  
  # model 10  forecast monthly data rolled into yearly and forecasted -- HW
  fit10 <- hw(rolling.yearly.sum,damped = TRUE)
  fcst10 <- forecast(fit10,h=12)
  fcst10$mean
  
  
  ###printing all forecast
  
  plot(test.ts)
  lines(test.ts,col=2)
  lines(window(fcst.avg,start =test.start, end = test.end),col=3)
  lines(window(fcst.mon.avg,start =test.start, end = test.end),col=4)
  lines(window(fcst.arima.1$mean,start =test.start, end = test.end),col=5)
  lines(window(fcst.stl$mean,start =test.start, end = test.end),col=6)
  lines(window(fcst4$mean,start =test.start, end = test.end),col=7)
  lines(window(fcst5.3$mean,start =test.start, end = test.end),col=8)
  legend("topright",lty=1, pch=2, col=2:8, cex = 0.5,
         c("Test Fcst","Average" ,"Monthly Average","Arima","STL","NNetar","ETS"))
  
  
  
  results.df <- data.frame( "Material" = c(material),
                            "Avg.fcst.rmse" =c(fcst.all.avg.rmse),
                            "Avg.monthly.fcst.rmse" = c(fcst.month.rmse),
                            "Arima.rmse" = c(arima.rmse),
                            "STL.rmse" = c(stl.rmse),
                            "Nnet.rmse" = c(nnet.rmse),
                            "Total.test.fcst" = c(sum(test.ts)),
                            "Total.avg.fcst" = c(sum(fcst.avg)),
                            "Total.month.avg.fcst" = c(sum(fcst.mon.avg)),
                            "Total.Arima.fcst" = c(sum(fcst.arima.1$mean)),
                            "Total.STL.fcst" = c(sum(fcst.stl$mean)),
                            "Total.Nnet.fcst" = c(sum(fcst4$mean)),
                            "Total.HW.Damped.Fcst" = c(sum(fcst5.1$mean)),
                            "Total.HW.Fcst" = c(sum(fcst5.2$mean)),
                            "Total.ets.Fcst" = c(sum(fcst5.3$mean)),
                            "Total.ets.damped.Fcst" = c(sum(fcst5.4$mean)),
                            "Total.Batch.Arima.Fcst" =  c(fcst6$mean[test.length]),
                            "Total.Batch.STL.Fcst"  = c(fcst7$mean[test.length]),
                            "Total.Batch.HW.Fcst" = c(fcst8$mean[test.length]),
                            "Total.Batch.ETS.Fcst" = c(fcst9$mean[test.length]),
                            "Total.Batch.HW.Damped.Fcst" = c(fcst10$mean[test.length]))
  
  
  
  results.df
  
}

```

````{r calling.generate.forecast}

#res <- generate.forecast(487999,c(2011,1),c(2015,1),c(2015,12))
#res <- generate.forecast(242061)
#melt(res)

#generating for all items

all.result.df  <- generate.forecast(material.input <- unique(trainFinal$Material)[1]) # to initialize

material.input <- unique(trainFinal$Material)[2:513] #change 20 for generating forecast for all items
for (mat in material.input ) { 

all.result.df <- rbind(all.result.df, generate.forecast(mat))

}

```

```{r accuracy}
all.result.df$MAE <- rep(1,nrow(all.result.df))
all.result.df$best.algo <- rep("",nrow(all.result.df))

for (i in 1:nrow(all.result.df)) {
  total.fcsts <- all.result.df[i,8:ncol(all.result.df)]
  test.fcst <- all.result.df[i,]$Total.test.fcst
  total.fcsts.mae <- abs(all.result.df[i,8:(ncol(all.result.df)-2)] - test.fcst)
  min.col <- apply(total.fcsts.mae,1,which.min)
  min.val <- total.fcsts.mae[min.col]
  all.result.df[i,]$best.algo <- colnames(total.fcsts.mae)[min.col]
  all.result.df[i,]$MAE <- min.val
  
}

write.csv(all.result.df, "allresult.csv")

```

#Connecting Tableau and R
```{r}
#install a package Rserve.
install.packages("Rserve")

#start the Rserve
library(Rserve)
Rserve(arg="--no-save")
